---
title: "mooreaviz"
author: "Jake Eisaguirre"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(raster)
library(viridis)
library(sf)
library(gstat)
library(sp)
library(leaflet)
library(automap)
library(shiny)
library(testthat)
library(htmlwidgets)
library(ggvoronoi)
library(car)
library(dismo)

```

# read in data
```{r}
n_data <- read_csv(here("knb-lter-mcr.5038.10", "MCR_LTER_Adam_EcolAp_N_summary_20200331.csv")) %>% 
  clean_names()

new_reef <- read_csv(here("reef_bound", "reef_bound.csv"))

bleach <- read_csv(here("MCR_LTER_Burkepile_Adams_2016_Coral_Bleaching_Survey_20191023.csv")) %>% 
  clean_names()
```

# Tidy format nitrogen data
```{r}

n_data <- n_data %>% 
  pivot_longer(!1:5, names_to = "type", values_to = "percent_n") %>% 
  separate(type, into = c("method","random", "date"), sep = "_") %>% 
  dplyr::select(-random)

```

# Prepping shapefiles, boundary, and gridded boundary
```{r}
#moorea crs EPSG:2976 
crs <- 2976

#reef bound in UTM
bound <- new_reef %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% 
  mutate(geometry = st_transform(geometry, "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs)) %>% 
  summarise(geometry = st_combine(geometry)) %>% 
  st_cast("POLYGON") %>% 
  st_set_crs(crs)

#Island Bound in UTM
island_bound_shp <- read_sf(here("island", "bh400kc3500.shp")) %>% 
  mutate(geometry = st_transform(geometry, "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs)) %>% 
  dplyr::select(geometry)

#clipping reef boundary 
island_bound <- st_crop(island_bound_shp, bound)

clipped_bound <- st_difference(bound,island_bound)


#making a gridded box with the bounds of the lat long in UTM 
grd_sf <- clipped_bound %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_make_grid(cellsize = c(100, 100), what = "centers") %>%
  st_as_sf() %>% 
  cbind(., st_coordinates(.)) 


#Clipping the gridded box by the reef bound
bound_grid <- st_intersection(grd_sf, clipped_bound)

#converting between formats and confirming grid
grd_sp <- as(bound_grid, "Spatial") 
gridded(grd_sp) <- TRUE           
grd_sp <- as(grd_sp, "SpatialPixels")


```

# Kriging january nitrogen data for Moorea lagoon/reef boundary
```{r}
# selecting n-jan data
jan_np_data <- n_data %>% 
  filter(date == "jan", method == "percent") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting jan data to sf and changing lat long to UTM
sf_jan_np_data <- st_as_sf(jan_np_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
jan_np_vario <- variogram(percent_n~1, as(sf_jan_np_data, "Spatial"))

plot(jan_np_vario)

v_mod_full_jan <- autofitVariogram(percent_n~1, as(sf_jan_np_data, "Spatial"))

v_mod_jan <- v_mod_full_jan$var_model

plot(v_mod_full_jan)


#krige 
jan_krig_p <- krige(percent_n~1, as(sf_jan_np_data, "Spatial"), 
                  grd_sp, model = v_mod_jan)

#convert kriged data to raster  
jan_ras <- raster(jan_krig_p)

```


# Kriging may nitrogen data for Moorea lagoon/reef boundary
```{r}
# selecting n-may data
may_np_data <- n_data %>% 
  filter(date == "may", method == "percent") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting may data to sf and changing lat long to UTM
sf_may_np_data <- st_as_sf(may_np_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
may_np_vario <- variogram(percent_n~1, as(sf_may_np_data, "Spatial"))

plot(may_np_vario)

v_mod_full_may <- autofitVariogram(percent_n~1, as(sf_may_np_data, "Spatial"))

v_mod_may <- v_mod_full_may$var_model

plot(v_mod_full_may)


#krige 
may_krig_p <- krige(percent_n~1, as(sf_may_np_data, "Spatial"), 
                  grd_sp, model = v_mod_may)

#convert kriged data to raster  
may_ras <- raster(may_krig_p)

plot(may_ras)
```


# Kriging July nitrogen data for Moorea lagoon/reef boundary

```{r}
# selecting n-july data
july_np_data <- n_data %>% 
  filter(date == "july", method == "percent") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting july data to sf and changing lat long to UTM
sf_july_np_data <- st_as_sf(july_np_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
july_np_vario <- variogram(percent_n~1, as(sf_july_np_data, "Spatial"))

plot(july_np_vario)

v_mod_full_july <- autofitVariogram(percent_n~1, as(sf_july_np_data, "Spatial"))

v_mod_july <- v_mod_full_july$var_model

plot(v_mod_full_july)


#krige 
july_krig_p <- krige(percent_n~1, as(sf_july_np_data, "Spatial"), 
                  grd_sp, model = v_mod_july)

#convert kriged data to raster  
july_ras <- raster(july_krig_p)

plot(july_ras)
```

# Kriging january nitrogen isotopic data for Moorea lagoon/reef boundary
```{r}
# selecting n-jan data
jan_ni_data <- n_data %>% 
  filter(date == "jan", method == "d15n") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting jan data to sf and changing lat long to UTM
sf_jan_ni_data <- st_as_sf(jan_ni_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
jan_ni_vario <- variogram(percent_n~1, as(sf_jan_ni_data, "Spatial"))

plot(jan_ni_vario)

v_mod_full_jan_i <- autofitVariogram(percent_n~1, as(sf_jan_ni_data, "Spatial"))

v_mod_jan_i <- v_mod_full_jan_i$var_model

plot(v_mod_full_jan_i)


#krige 
jan_krig_i <- krige(percent_n~1, as(sf_jan_ni_data, "Spatial"), 
                  grd_sp, model = v_mod_jan_i)

#convert kriged data to raster  
jan_ras_i <- raster(jan_krig_i)

plot(jan_ras_i)
```
# Kriging may nitrogen isotopic data for Moorea lagoon/reef boundary
```{r}
# selecting n-may data
may_ni_data <- n_data %>% 
  filter(date == "may", method == "d15n") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting may data to sf and changing lat long to UTM
sf_may_ni_data <- st_as_sf(may_ni_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
may_ni_vario <- variogram(percent_n~1, as(sf_may_ni_data, "Spatial"))

plot(may_ni_vario)

v_mod_full_may_i <- autofitVariogram(percent_n~1, as(sf_may_ni_data, "Spatial"))

v_mod_may_i <- v_mod_full_may_i$var_model

plot(v_mod_full_may_i)


#krige 
may_krig_i <- krige(percent_n~1, as(sf_may_ni_data, "Spatial"), 
                  grd_sp, model = v_mod_may_i)

#convert kriged data to raster  
may_ras_i <- raster(may_krig_i)

plot(may_ras_i)
```


# Kriging july nitrogen isotopic data for Moorea lagoon/reef boundary
```{r}
# selecting n-july data
july_ni_data <- n_data %>% 
  filter(date == "july", method == "d15n") %>%
  dplyr::select(longitude, latitude, percent_n) %>% 
  na.omit()


#converting july data to sf and changing lat long to UTM
sf_july_ni_data <- st_as_sf(july_ni_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs))


#vario gram and model selection 
july_ni_vario <- variogram(percent_n~1, as(sf_july_ni_data, "Spatial"))

plot(july_ni_vario)

v_mod_full_july_i <- autofitVariogram(percent_n~1, as(sf_july_ni_data, "Spatial"))

v_mod_july_i <- v_mod_full_july_i$var_model

plot(v_mod_full_july_i)


#krige 
july_krig_i <- krige(percent_n~1, as(sf_july_ni_data, "Spatial"), 
                  grd_sp, model = v_mod_july_i)

#convert kriged data to raster  
july_ras_i <- raster(july_krig_i)

plot(july_ras_i)
```


# Jakes Box
```{r}

```


# Percent Bleached
```{r}
# #Select column
# bleaching_data <- bleach %>% 
#   dplyr::select(latitude, longitude, percent_bleached) %>% 
#   na.omit()
# 
# #converting bleach data to sf and changing lat long to UTM
# sf_bleached_data <- st_as_sf(bleaching_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
#   cbind(st_coordinates(.)) %>% 
#   mutate(geometry = st_transform(geometry, 
#                                  "+proj=utm +zone=6 +south +ellps=intl 
#                                  +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
#                                  crs = crs))
# 
# #vario gram and model selection 
# bleach_vario <- variogram(percent_bleached~1, as(sf_bleached_data, "Spatial"))
# 
# plot(bleach_vario)
# 
# v_mod_full_bleach <- autofitVariogram(percent_bleached~1, as(sf_bleached_data, "Spatial"))
# 
# v_mod_bleach <- v_mod_full_bleach$var_model
# 
# plot(v_mod_full_bleach)
# 
# july_krig_p <- krige(percent_bleached~1, as(sf_bleached_data, "Spatial"), 
#                   grd_sp, model = "power")

```


# rasters and palettes 
```{r}
# percent nitrogen 
jan_data <- as.data.frame(rasterToPoints(jan_ras))
pal_jan <- colorNumeric(palette = viridis((25), option = "plasma"), domain = jan_data$var1.pred)

may_data <- as.data.frame(rasterToPoints(may_ras))
pal_may <- colorNumeric(palette = viridis((25), option = "plasma"), domain = may_data$var1.pred)

july_data <- as.data.frame(rasterToPoints(july_ras))
pal_july <- colorNumeric(palette = viridis((25), option = "plasma"), domain = july_data$var1.pred)

# isotopic nitrogen 

jan_i_data <- as.data.frame(rasterToPoints(jan_ras_i))
pal_jan_i <- colorNumeric(palette = viridis((25), option = "plasma"), domain = jan_data$var1.pred)

may_i_data <- as.data.frame(rasterToPoints(may_ras_i))
pal_may_i <- colorNumeric(palette = viridis((25), option = "plasma"), domain = may_data$var1.pred)

july_i_data <- as.data.frame(rasterToPoints(july_ras_i))
pal_july_i <- colorNumeric(palette = viridis((25), option = "plasma"), domain = july_data$var1.pred)

```



# leaf for slides
```{r}

leaflet() %>% 
  addProviderTiles("Esri.WorldImagery") %>% 
  addCircleMarkers(lng = jan_np_data$longitude, lat = jan_np_data$latitude,
                   color = "black", group = "Observations") %>% 
  addRasterImage(jan_ras, colors = "plasma", group = "January N", opacity = 0.7) %>% 
  addRasterImage(may_ras, colors = "plasma", group = "May N", opacity = 0.7) %>% 
  addRasterImage(july_ras, colors = "plasma", group = "July N", opacity = 0.7) %>% 
  addRasterImage(jan_ras_i, colors = "plasma", group = "January N Isotopic", opacity = 0.7) %>% 
  addRasterImage(may_ras_i, colors = "plasma", group = "May N Isotopic", opacity = 0.7) %>% 
  addRasterImage(july_ras_i, colors = "plasma", group = "July N Isotopic", opacity = 0.7) %>%
  addLayersControl(
    baseGroups = c("January N", 
                   "May N", 
                   "July N", 
                   "January N Isotopic", 
                   "May N Isotopic", 
                   "July N Isotopic"),
    overlayGroups = c("Observations"),
    options = layersControlOptions(collapsed = FALSE))  %>% 
  addLegend(data = jan_data, title = 'January N', pal = pal_jan,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "January N") %>% 
  onRender("
    function(el, x) {
      var updateLegend = function () {
          var selectedGroup =       document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);

          document.querySelectorAll('.legend').forEach(a => a.hidden=true);
          document.querySelectorAll('.legend').forEach(l => {
            if (l.children[0].children[0].innerText == selectedGroup) l.hidden=false;
          });
      };
      updateLegend();
      this.on('baselayerchange', e => updateLegend());
    }") %>% 
  addLegend(data = may_data, title = 'May N', pal = pal_may,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "May N") %>% 
  addLegend(data = may_data, title = 'July N', pal = pal_may,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "July N") %>% 
  addLegend(data = may_data, title = 'January N Isotopic', pal = pal_may,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "January N Isotopic") %>% 
  addLegend(data = may_data, title = 'May N Isotopic', pal = pal_may,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "May N Isotopic") %>% 
  addLegend(data = may_data, title = 'July N Isotopic', pal = pal_may,
            position = "bottomright", 
            values = ~var1.pred, opacity = 1, group = "July N Isotopic")

#saveWidget(leaf, file = here("leaf.html"))
```

# random ggplot for slides
```{r}


ggplot(data = jan_data) + 
  geom_raster(aes(x = x, y = y, fill = var1.pred)) +
  scale_fill_viridis_c(option = "plasma", "Predicted N") +
  theme_classic() +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  ggtitle("January Turbinaria Ornata Nitrogen")

ggsave(here("pred_n.jpeg"), dpi = 300, width = 7, height = 4, units = "in")


```
#### Model Check
```{r}

#### Kriging can be understood as a two-step process: first, the spatial covariance structure of the sampled points is determined by fitting a variogram; and second, weights derived from this covariance structure are used to interpolate values for unsampled points or blocks across the spatial field.

#### variogram (sometimes called a “semivariogram”) is a visual depiction of the covariance exhibited between each pair of points in the sampled data. For each pair of points in the sampled data, the gamma-value or “semi-variance” (a measure of the half mean-squared difference between their values) is plotted against the distance, or “lag”, between them.

###Ordinary kriging, for which the assumption of stationarity (that the mean and variance of the values is constant across the spatial field) must be assumed. This is one of the simplest forms of kriging, but the stationarity assumption is not often met in applications relevant to environmental health, such as air pollution distributions.

#### Universal kriging is used to estimate spatial means when the data have a strong trend and the trend can be modeled by simple functions. Trend is scale dependent.

####Ordinary Kriging is a spatial estimation method where the error variance is minimized. This error variance is called the kriging variance. It is based on the configuration of the data and on the variogram, hence is is homoescedastic

#normally distributed data points
ggplot(jan_n_data) +
  geom_histogram(aes(x = percent_n)) +
  theme_classic()
ggsave(here("norm_dist.jpeg"), dpi = 300)

qqPlot(jan_n_data$percent_n)

#data is stationary 

spdf <- as(clipped_bound, "Spatial")

utm_data <- st_as_sf(jan_n_data, coords = c('longitude', 'latitude'), crs = 4326) %>% 
  cbind(st_coordinates(.)) %>% 
  mutate(geometry = st_transform(geometry, 
                                 "+proj=utm +zone=6 +south +ellps=intl 
                                 +towgs84=162,117,154,0,0,0,0 +units=m +no_defs", 
                                 crs = crs)) %>% 
  as.data.frame(dplyr::select(percent_n, geometry))

ggplot(jan_n_data) +
  geom_voronoi(aes(x = longitude, y = latitude, fill = percent_n), outline = spdf) +
  scale_fill_viridis_c(option = "plasma", "Predicted N")
```
### This is all exploratory stuff and writing .shp for reef bound
```{r}
#moorea crs EPSG:2976???
crs <- 2976

jan_n_data <- n_data %>% 
  dplyr::select(longitude, latitude, percent_n_jan) %>% 
  na.omit()


#data plot
ggplot() +
  geom_point(data = jan_n_data, mapping = aes(x = longitude, y = latitude, color = percent_n_jan),
             size = 3) + 
  scale_color_viridis(option = "B") +
  theme_classic()


#this boundary is normal lat long
bound_explore <- nut_bound %>% 
  st_as_sf(coords = c("X", "Y"), crs = crs) %>%
  summarise(geometry = st_combine(geometry)) %>% 
  st_cast("POLYGON") %>% 
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84"))

#code to save shape file. once we determine best shape file, current one is wrong. might be better to get shape file of moorea
#and use that for inner bound and then use current outter bound.
shp_bound <- st_write(bound_explore, here("shape_files", "bound.shp"), delete_dsn=TRUE)

bound_explore <- read_sf(here("shape_files", "bound.shp")) %>% 
   mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  st_set_crs(crs)


#reef bound and data point plot
ggplot()+
  geom_sf(data = bound_explore, col = "black", fill = NA) +
  theme_classic() +
  geom_point(data = jan_n_data, mapping = aes(x = longitude, y = latitude, color = percent_n_jan),
             size = 1) + 
  scale_color_viridis(option = "B", "Predicted N") +
  coord_sf(crs = crs) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  ggtitle("January Turbinaria Ornata Nitrogen")

ggsave(here("explore_pred_n.jpeg"), dpi = 300, width = 7, height = 4, units = "in")



```
